---
title: "Assessment"
author: "Freya"
date: "2024-11-01"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assessing opioid replacement therapies by region 

Justified – understanding of the data and their context 
Introduction 
Key results 
Conclusion
References (for reason chosen research question)
Narrative shows critical thinking
Multiple data sources 
Recommendations and conclusions are grounded in the data 
Limitations of the data set discussed. Next steps suggested in terms of data that would allow for further analysis. 

## Introduction 

My project will analyse X, Y, Z 

```{r cars}
library(ggplot2)
summary(cars)
```

## Key Results 

>4 plots and / or tables 
No use of settings other than default 
Labelled, titled plots 
Labelled titled tables 
Appropriate and thoughtful data visualisation with some use of non-default settings (gt)
Faceting, multilayered plots, interactive visualisation 

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

## Conclusion 

### Recommendations from analysis 

### Limitations of the dataset and suggestions for future analysis


## References 


