---
title: "Assessment"
author: "B273025"
date: "`r Sys.Date()`"
bibliography: "../zotero_citations.json"
link-citations: true
output: 
  html_document: 
    code_download: true
    toc: true
    theme: flatly

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache=FALSE)
```

# Emergency Contraceptive Pill and Sexually Transmitted Infections

Mark scheme:
Justified – understanding of the data and their context  
Introduction  
Key results   
Conclusion 
References (for reason chosen research question)  
Narrative shows critical thinking  
Multiple data sources   
Recommendations and conclusions are grounded in the data  
Limitations of the data set discussed. Next steps suggested in terms of data that would allow for further analysis.   

## Introduction 

There are two types of emergency contraceptive pill prescribed by pharmacies.

>* **Levonogestrel** (brand name: *Levonelle*) taken within three days of unprotected sex
>* **Ulipristal Acetate** (brand name: *ellaOne*) taken within five days of unprotected sex  


This report explores 

1) Prescribing patterns of emergency contraceptive pill (ECP) between 2019 to 2023. 

2) How ECP and antibiotics for STI prescriptions differ in geography comparing university town to a non-university town 

3) Compare patterns in ECP and STI antibiotics prescribing during Covid-19 lockdowns - did the need for ECP or incidence of STI antibiotics prescription go down during lockdowns?

4) How does SIMD relate to prescription of ECP / antibiotics for STIs 

```{r}
# load necessary libraries
library(tidyverse)
library(janitor) 
library(gt) 
library(here)
library(lubridate)
library(patchwork)
```

I have chosen to look at all prescriptions from 2020 and 2023. To do this I downloaded the prescriptions data for each month from: https://www.opendata.nhs.scot/dataset/prescriptions-in-the-community  

I wanted to explore how many ECP and antibiotics for STIs were prescibed during across 2023.



>4 plots and / or tables 
No use of settings other than default 
Labelled, titled plots 
Labelled titled tables 
Appropriate and thoughtful data visualisation with some use of non-default settings (gt)
Faceting, multilayered plots, interactive visualisation 

Read in data:
```{r}
# Read in the Health Board names (HB_names) and clean their column names. I have done this at the beginning to make the datasets globally available  for the rest of my code. 
HB_names <- read_csv("https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/652ff726-e676-4a20-abda-435b98dd7bdc/download/hb14_hb19.csv") %>% 
  clean_names() # standardises the column names by making all names lower case and underscores for spaces. This makes code consistent and easier to manipulate.

# Read in female population data per Health Board by using the most recent census data. 
population_data <- read_csv(here("data", "UV103_age_health_board_census.csv"), skip = 10) %>% # locates the csv and excludes the first 10 lines in the csv as they are redundant
  rename(Spare = "...6", # remove unused columns
         hb_name = "Health Board Area 2019",
         hb_population = Count) %>% # rename() formats the data to match the prescriptions dataframe
  filter(Age == "All people" & Sex == "Female") %>% # filter female population, as men do not take ECP
  select(hb_name, hb_population) %>% 
  mutate(hb_name = paste("NHS", hb_name)) %>%  # change hb_name column format to match the Health Board dataframe format
  clean_names() # ensures consistency
```

Define a function to read in the prescription data for a defined year between 2019 and 2022:
```{r}
read_all_prescriptions <- function(year){
  all_files <- list.files(here("data", paste0("all_months_", year)), pattern = "csv") # I downloaded 12 months of prescription data and placed them within folders named all_months_year. list.files() identifies the relevant monthly prescription data for the year in the function.
  
  all_prescriptions <- all_files %>% 
    map_dfr(~read_csv(here("data", paste0("all_months_", year),.))) %>%
    clean_names() %>% 
    drop_na(bnf_item_description) # drop the rows with missing values 
  return(all_prescriptions)
}
```


```{r}

# I have used the function read_all_prescriptions() to read in prescription data from 2019 to 2022.
all_prescriptions_2019 <- read_all_prescriptions(2019)
all_prescriptions_2020 <- read_all_prescriptions(2020)
all_prescriptions_2021 <- read_all_prescriptions(2021)
all_prescriptions_2022 <- read_all_prescriptions(2022)

# 2019 dataset has hbt2014 as a column name instead of hbt. Rename to make column name consistent
all_prescriptions_2019 <- all_prescriptions_2019 %>% 
  rename(hbt = "hbt2014")

# combine 4 years of data into one dataset to make it easier to wrangle
combined_prescriptions <- bind_rows(
  all_prescriptions_2019,
  all_prescriptions_2020,
  all_prescriptions_2021,
  all_prescriptions_2022
)

# select emergency contraceptive pill (ECP) to make dataframe smaller before continuing with data wrangling and joining datasets
ECP_prescriptions <- combined_prescriptions %>% 
  mutate(
    date = parse_date_time(paid_date_month, "ym"), # use lubridate to format date
    drug_simple = case_when(
      str_detect(bnf_item_description, "LEVONO") ~ "Levonorgestrel",
      str_detect(bnf_item_description, "ULIPR") ~ "Ulipristal Acetate",
      TRUE ~ "Other")) %>% # each drug has different dosages so I used case_when() to group them by their drug names 
  filter(drug_simple != "Other") %>% # remove other prescriptions from the dataset 
  group_by(hbt, date, drug = drug_simple) %>%  
  summarise(total_quantity = sum(paid_quantity, na.rm = TRUE)) # sum prescriptions of ECP

# full join HB_names and population_data to all_prescriptions dataset
ECP_prescriptions_population <- ECP_prescriptions %>%
  full_join(HB_names, by = c("hbt" = "hb")) %>%          # Join with Health Board names
  full_join(population_data, by = "hb_name") %>%         # Join with population data
  select(hb_name, date, drug, total_quantity,hb_population) %>% # Select columns of interest
  drop_na(total_quantity) %>%                                          # Drop rows with missing values
  clean_names()  

```


## Key Results 

### Seasonal trends in prescribing of Emergency Contracpetion 

This graph explores seasonal trends in prescribing of the two most commonly prescribed emergency contraceptive pills, Levonogestrel and Ulipristal Acetate, between 2019 to 2022 by Health Board. I plotted this to see if events such as Valentines day, University term start dates, or national holidays effect emergency contraceptive prescribing rates. I included 4 years of consecutive data to see if a one-time event, such as the Common Wealth Games caused any spikes in emergency contraception prescribing. Additionally by plotting the 4 consecutive years we can see if Covid-19 lockdowns in the years 2020 and 2021 affected emergency contraception prescribing rates.

Plot:
```{r}
#
ECP_prescriptions_population_rate <- ECP_prescriptions_population %>% 
  mutate(prescriptions_per_10000 = (total_quantity / hb_population)*10000) %>% # calculate the ECP prescription rate
  mutate(month = factor(month(date), levels = 1:12, labels = month.abb), .after = date) %>%   # extract month as a factor with labels to help when plotting
  mutate(year = factor(year(date)), .after = month) %>%  # extract year as a factor with labels to help when plotting
  ungroup()


ECP_plot <- ECP_prescriptions_population_rate %>%
  group_by(year, month, drug) %>% 
  summarise(prescriptions_per_10000 = sum(prescriptions_per_10000)) %>% 
  ggplot(aes(x = month, y = prescriptions_per_10000, color = drug, group = drug)) +
  geom_line() +
  facet_wrap(~year)+
  labs(
    title = "Line graphs to show the seasonal trends in \ntotal Emergency Contraception Prescriptions for each year from 2019 to 2022",#line break for formating
    x = "Month",
    y = "Prescriptions per 10,000 poeple",
    color = "Type of Emergency \nContraception prescribed" #show differences in type of pill prescribed
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 12, hjust = 0.5),  # Title size and centre alignment
        axis.title.x = element_text(size = 10),  # X-axis title size
        axis.title.y = element_text(size = 10), # Y-axis title size
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),# X-axis text size and angle
        axis.text.y = element_text(size = 8),  # Axis text size
        strip.text = element_text(size = 10),  # Facet label text size
        legend.title = element_text(size = 10),  # Legend title size
        legend.text = element_text(size = 8),  # Legend text size
        legend.position = "bottom")

ECP_plot

```

Trends to note:

* Levonorgestrel is more commonly prescribed than Ulipristal Acetate. 
* 2019 and 2022 represent years without Covid-19 lockdown measures. 2019 and 2022 have a peak in prescriptions of Levonorgestrel in October and August respectively. This perhaps reflects increased sexual activity coupled with the return to university with 'fresher' week. 
* Between Febuary and May of 2020 the prescription rate of Levonorgestrel decreases. This is likely due to decreased sexual activity and risk taking behaviours due to the stringent national 'lockdown' measures introduced between X to Y in response to the Covid-19 pandemic. 
* At the start of 2021 prescription rate of Levonorgestrel is very low. Likely due to additional Covid-19 lockdowns.

### Impact of deprivation on the type of emergency contraceptive prescribed

How does deprivation influence type of contraception prescribed?
I wanted to explore if there was any correlation between the type of contraception being prescribed and deprivation. To do this I made a ratio of levonogestrel to total emergency contraception prescriptions. 

> A/(A+B)
> Where A = Levonogestrel (prescribed within 3 days of unprotected sex)
> B = Ulipristal Acetate (prescribed within 5 days of unprotected sex)

This is important to identify any variation in prescribing patterns in more deprived areas. Differences may suggest patients access health services later, so have to use Ulipristal Acetate, or that GPs or Pharmacists have differences in prescribing preferences in more deprived areas.

To measure deprivation I have used the Scottish Index of Multiple Deprivation. I took the SIMD 2020v2 dataset from Public Health Scotland website [available from: https://www.opendata.nhs.scot/gl/dataset/scottish-index-of-multiple-deprivation/resource/acade396-8430-4b34-895a-b3e757fa346e ] as this dataset contains SIMD decile weighted per Health Board population.

When interpreting SIMD in deciles rank 1 is considered the most deprived, and rank 10 is least deprived. 

```{r}
# # read in html
# SIMD <- read_csv("https://www.opendata.nhs.scot/gl/dataset/78d41fa9-1a62-4f7b-9edb-3e8522a93378/resource/acade396-8430-4b34-895a-b3e757fa346e/download/simd2020v2_22062020.csv") %>%
#   clean_names() %>%
#   select(data_zone, simd2020v2hb_decile)
# 
# gp_addresses <- read_csv("https://www.opendata.nhs.scot/dataset/f23655c3-6e23-4103-a511-a80d998adb90/resource/9c1dccc7-7632-4b13-b451-092bd57973a4/download/practice_contactdetails_apr2023-open-data-1.csv") %>% 
#   clean_names() %>% 
#   select(practice_code, gp_practice_name, data_zone)
# 
# data_zones <- read_csv("https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/395476ab-0720-4740-be07-ff4467141352/download/dz2011_codes_and_labels_21042020.csv") %>% 
#   clean_names() %>% 
#   select(data_zone, data_zone_name)
# 
# # Create ECP_prescriptions_GP_location by using the GP_addresses dataset to map the GP practice code to a datazone. I could then use datazone to full_join() the SIMD dataset to the prescriptions dataset. 
# 
# ECP_prescriptions_GP_location <- combined_prescriptions %>%
#   mutate(
#     date = parse_date_time(paid_date_month, "ym"),
#     drug_simple = case_when(
#       str_detect(bnf_item_description, "LEVONO") ~ "Levonorgestrel",
#       str_detect(bnf_item_description, "ULIPR") ~ "Ulipristal Acetate",
#       TRUE ~ "Other")) %>% 
#   filter(drug_simple != "Other") %>%
#   group_by(gp_practice, date, drug = drug_simple) %>% # this time keeping in gp_practice
#   summarise(total_quantity = sum(paid_quantity, na.rm = TRUE)) %>%
#   full_join(HB_names, by = c("hbt" = "hb")) %>% # Join with Health Board names
#   full_join(population_data, by = "hb_name") %>% # Join with population data
#   mutate(prescriptions_per_10000 = (total_quantity / hb_population)*10000) %>% 
#   left_join(gp_addresses, by = c("gp_practice" = "practice_code")) %>% 
#   #left_join(data_zones, by = "data_zone") %>% 
#   left_join(SIMD, by = "data_zone") %>%
#   ungroup() %>%
#   drop_na() %>% # Drop rows with missing values
#   clean_names()
# 
# ECP_SIMD_barchart <- ECP_prescriptions_GP_location %>% 
#   group_by(SIMD2020v2hb_decile, drug) %>% 
#   summarise(total_prescriptions_per_10000 = sum(prescriptions_per_10000, na.rm = TRUE)) %>%  # summarise the prescriptions per 10,000 
#   ungroup() %>%
#   ggplot(aes(x = total_prescriptions_per_10000, y = factor(simd2020v2hb_decile, levels = 1:10), fill = drug)) +
#   geom_bar() +
#   theme(legend.position = "bottom")
# 
# ECP_SIMD_barchart
```

Prescribing choices vary per region -  




Data sets to consider exploring: 

```{r}
# Data sets to consider exploring: 
#data_zones_2011 <- read_csv("https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/395476ab-0720-4740-be07-ff4467141352/download/dz2011_codes_and_labels_21042020.csv") %>% 
#clean_names()

#SIMD_2020 <- read_csv(here("data", "scottish-index-of-multiple-deprivation.csv")) %>%
#clean_names()

#gp_addresses_2020 <- read_csv("https://www.opendata.nhs.scot/dataset/f23655c3-6e23-4103-a511-a80d998adb90/resource/42391720-7dcb-48a2-8070-b9d63b246ac6/download/practice_contactdetails_oct2020-open-data.csv") %>% 
#clean_names() %>% 
#select(practice_code, gp_practice_name, data_zone)

#data_zones <- read_csv("https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/395476ab-0720-4740-be07-ff4467141352/download/dz2011_codes_and_labels_21042020.csv") %>% 
#clean_names() %>% 
#select(data_zone, data_zone_name)
```


## Conclusion 

### Recommendations from analysis 

### Limitations of the dataset and suggestions for future analysis

antibiotics are prescribed for pneumonia / other infections. Not specific to STIs. Maybe find what percentage of antibiotics for doxy / azithro are prescribed for STIs in scotland and use this as an assumption when completing further analysis?



## References 
I used https://gsverhoeven.github.io/post/zotero-rmarkdown-csl/ to set up citations. Chosen BMJ style as common. 

