---
title: "Assessment"
author: "B273025"
date: "`r Sys.Date()`"
bibliography: "../zotero_citations.json"
link-citations: true
output: 
  html_document: 
    code_download: true
    toc: true
    theme: flatly

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Emergency Contraceptive Pill and Sexually Transmitted Infections

Justified – understanding of the data and their context  
Introduction  
Key results   
Conclusion 
References (for reason chosen research question)  
Narrative shows critical thinking  
Multiple data sources   
Recommendations and conclusions are grounded in the data  
Limitations of the data set discussed. Next steps suggested in terms of data that would allow for further analysis.   

## Introduction 

There are two types of emergency contraceptive pill prescribed by pharmacies.

>* **Levonogestrel** (brand name: *Levonelle*) taken within three days of unprotected sex
>* **Ulipristal Acetate** (brand name: *ellaOne*) taken within five days of unprotected sex  

STIs pose a  significant public health, with recent trends showing an increase in prevalence. 
The two most common sexually transmitted infections in Scotland are 
>* **chlamydia**, treated with *doxycylcine* or *azithromycin*  
>* **gonorrhea**, treated with *intramuscular ceftriaxone*.  


This report explores 

1) Prescribing patterns of oral contraceptive pill and antibiotics across 2023 (any time trends eg University return dates, festivals), considering if the prescription of OCP mirrors number of prescriptions of antibiotics used to treat STIs

2) How OCP and antibiotics for STI prescriptions differ in geography comparing university town to a non-university town 

3) Compare patterns in OCP and STI antibiotics prescribing during Covid-19 lockdowns - did the need for OCP or incidence of STI antibiotics prescription go down during lockdowns?

```{r}
library(tidyverse)
library(janitor) 
library(gt) 
library(here)
```

I have chosen to look at all prescriptions from 2020 and 2023. To do this I downloaded the prescriptions data for each month from: https://www.opendata.nhs.scot/dataset/prescriptions-in-the-community  

I wanted to explore how many OCP and antibiotics for STIs were prescibed during across 2023.

```{r}
#downloaded consecutive data for prescriptions for each month of 2020 and each month of 2023 and placed them into a folder named all_months_2020 and all_months_2023 respectively. I then combined these files into one variable called all_prescriptions_2020 or all_prescriptions_2023.

all_files_2020 <- list.files(here("data","all_months_2020"), pattern = "csv")
all_prescriptions_2020 <- all_files_2020 %>% 
  map_dfr(~read_csv(here("data","all_months_2020", .))) %>% 
  clean_names()

all_files_2023 <- list.files(here("data","all_months_2023"), pattern = "csv")
all_prescriptions_2023 <- all_files_2023 %>% 
  map_dfr(~read_csv(here("data","all_months_2023", .),col_types = cols(DMDCode = col_character()))) %>% 
  clean_names()

#why did I need to use DMDCode col_character() here?

HB_names <- read_csv("https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/652ff726-e676-4a20-abda-435b98dd7bdc/download/hb14_hb19.csv") %>% 
  clean_names()

#full_join to add health board name to the prescriptions dataset

all_prescriptions_2020 <- all_prescriptions_2020 %>% 
  full_join(HB_names, by = c("hbt" = "hb")) %>% 
  select(hb_name, hbt:paid_date_month)


all_prescriptions_2023 <- all_prescriptions_2023 %>% 
  full_join(HB_names, by = c("hbt" = "hb")) %>% 
  select(hb_name, hbt:paid_date_month)

#data_zones_2011 <- read_csv("https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/395476ab-0720-4740-be07-ff4467141352/download/dz2011_codes_and_labels_21042020.csv") %>% 
  #clean_names()

#SIMD_2020 <- read_csv(here("data", "scottish-index-of-multiple-deprivation.csv")) %>%
  #clean_names()

#gp_addresses_2020 <- read_csv("https://www.opendata.nhs.scot/dataset/f23655c3-6e23-4103-a511-a80d998adb90/resource/42391720-7dcb-48a2-8070-b9d63b246ac6/download/practice_contactdetails_oct2020-open-data.csv") %>% 
  #clean_names() %>% 
  #select(practice_code, gp_practice_name, data_zone)

#data_zones <- read_csv("https://www.opendata.nhs.scot/dataset/9f942fdb-e59e-44f5-b534-d6e17229cc7b/resource/395476ab-0720-4740-be07-ff4467141352/download/dz2011_codes_and_labels_21042020.csv") %>% 
  #clean_names() %>% 
  #select(data_zone, data_zone_name)

# Use data from the most recent census to find the number of people in each healthboard 
population_data <- read_csv(here("data", "UV103_age_health_board_census.csv"), skip = 10) %>% 
  rename(Spare = "...6",
         hb_name = "Health Board Area 2019",
         hb_population = Count) %>% 
  filter(Age == "All people" & Sex == "All people") %>% 
  select(hb_name, hb_population) %>% 
  mutate(hb_name = paste("NHS", hb_name))

# join this to the 2023 prescriptions dataset 
all_prescriptions_2023 <- all_prescriptions_2023 %>% 
  full_join(population_data)

#initial exploration of dataset

OCP <- all_prescriptions_2023 %>% 
  filter(str_detect(bnf_item_description, "LEVONO|ULIPR"))
abx <- all_prescriptions_2023 %>% 
  filter(str_detect(bnf_item_description, "AZITHRO|DOXY|CEFTRI"))

OCP_abx_per_head <- all_prescriptions_2023 %>% 
  drop_na()%>% 
  filter(str_detect(bnf_item_description, "LEVONO|ULIPR")) %>% 
  group_by(hb_name) %>% 
  summarise(quantity_per_head = sum(paid_quantity)/mean(hb_population))

OCP_abx_per_head_chart <- OCP_abx_per_head %>% 
  ggplot(aes(x = quantity_per_head, y = reorder(hb_name, quantity_per_head))) +
  geom_col() +
  labs(title = "Prescription of oral combined contraceptive pill per person in each health board",
       x = "Quantity per head",
       y = "Health Board")
 
OCP_abx_per_head_chart

```


## Key Results 


>4 plots and / or tables 
No use of settings other than default 
Labelled, titled plots 
Labelled titled tables 
Appropriate and thoughtful data visualisation with some use of non-default settings (gt)
Faceting, multilayered plots, interactive visualisation 

You can also embed plots, for example:

```{r pressure, echo=FALSE}

```

## Conclusion 

### Recommendations from analysis 

### Limitations of the dataset and suggestions for future analysis


## References 
I used https://gsverhoeven.github.io/post/zotero-rmarkdown-csl/ to set up citations. Chosen BMJ style as common. 

